\texttt{apt-p2p}: A Peer-to-Peer Distribution System for Software Package Releases and Updates

### 2008-08-29 docs/paper/paper.tex
JC's updated changes to the INFOCOM paper.

\begin{abstract}
A large amount of free software packages are available over the
Internet from many different distributors. Most of these
distributors use the traditional client-server model to handle
requests from users. However, there is an excellent opportunity to
use peer-to-peer techniques to reduce the cost of much of this
distribution, especially due to the altruistic nature of many of
these users. There are no existing solutions suitable for this
situation, so we present a new technique for satisfying the needs of
this P2P distribution which is generally applicable to many of
these distributors' systems. Our method makes use of a DHT for
storing the location of peers, using the cryptographic hash of the
package as a key. To show the simplicity and functionality of our model, we
implement a solution for the distribution of Debian software
packages, including details on the DHT customizations and system optimizations needed. Finally, we
analyze our system to determine how it is performing and what effect
it is having.
\end{abstract}

@@

There are a large number of free software package distributors using
package distribution systems over the Internet to distribute
software to their users. These distributors have developed many
different methods for this distribution, but they almost exclusively
use a client-server model to satisfy user requests. The popularity
and number of users results in a large number of requests, which
usually requires a network of mirrors to handle. Due to the free
nature of this software, many users are willing and able to
contribute upload bandwidth to this distribution, but have no
current way to do this.

We present a new peer-to-peer distribution model to meet these
demands. It is based on many previous implementations of successful
peer-to-peer protocols, especially distributed hash tables (DHT) and
BitTorrent. The model relies on the pre-existence of cryptographic
hashes of the packages, which uniquely identify it for a
request from other peers. If the peer-to-peer download fails, then
the original request to the server is used as a fallback to prevent
any dissatisfaction from users. The peer can then share this new
package with others through the P2P system.

First, we examine the opportunity that is available for many of
these free software package distributors. We present an overview of
a system that would efficiently satisfy the demands of a large
number of users, and should significantly reduce the currently
substantial bandwidth requirements of hosting this software. We then
present an example implementation based on the Debian package
distribution system. This implementation will be used by a large
number of users, and serves as an example for other free software
distributors of the opportunity that can be met with such a system.
Finally, we analyze our currently deployed implementation to
determine how effective it is at meeting our goals, and to see what
effect it is having on the Debian package distribution system.

@@

The package management software that downloads packages displays
some kind of indication of speed and completeness for users to
watch. Since previous client-server downloads occurred in a sequential
fashion, the package management software also measures the speed
based on sequential downloading. This requires the peer-to-peer
solution to be reasonably responsive at retrieving packages
sequentially.

@@

On the other hand, there are aspects of BitTorrent that are no
longer needed in a system such as this one. With altruistic peers
and all files being available without uploading, incentives to share
become a less important issue. Also, the availability of seeds are
not critical either, as the mirror sites serve in that capacity
already.

@@

We assume that the user is attempting to download packages from a
server, using package management software to find and retrieve the
packages. Further, we assume that the requests from the user to the
server are able to be proxied by our peer-to-peer program, and that
the server is always available and has all of the package files.
Finally, the cryptographic hash of the packages must be available
separately from the package itself, and is usually contained in an
\emph{index} file which also contains all the packages' names,
locations and sizes.

@@

Our model for using P2P to enhance such a system is shown in
Figure~\ref{model}. As shown in Phase~1, our program will act as a
proxy (1,2), downloading (3) and caching all files communicated
between the user and the server (4). It will therefore also have
available the index files containing the cryptographic hashes all
packages. Later, in Phase~2, upon receiving a request from the user
to download a package (5), our program will search the index files
for the package being requested and find its hash (6). This hash can
then be looked up recursively in the DHT (7), which will return a
list of peers that have the package already (8). The package can
then be downloaded from the peers (11,12), it can be verified using
the hash (13), and if valid can be returned to the user (14). The
current nodes location is also added to the DHT for that hash (15),
as it is now a source for others to download from.

@@

There are several ways to implement the desired P2P functionality
into the existing package management software. The functionality can
be directly integrated through modifications to the software, though
this could be difficult as the P2P functionality should be running
at all times. This is needed both for efficient lookups and to
support uploading of already downloaded packages. Unfortunately, the
package management tools typically only run until the download and
install request is complete.

Many of the package management software implementations use HTTP
requests from web servers to download the packages, which makes it
possible to implement the P2P aspect as an almost standard HTTP
caching proxy. This proxy will run as a daemon in the background,
listening for requests from the package management tool for package
files, as well as serving (uploading) cached package to other peers.
It will get uncached requests first from the P2P system, or
fall back to the normal HTTP request from a server should it not
be found. For methods that don't use HTTP requests, other types of
proxies may also be possible.

@@

Although not necessary, we recommend implementing a download
protocol that is similar to the protocol used to fetch packages from
the distributor's servers. This simplifies the P2P program, as it
can then treat peers and mirrors almost identically when requesting
packages. In fact, the mirrors can be used when there are only a few
slow peers available for a file to help speed up the download
process.

@@

Since the package management system only stores a hash of the entire
package, and not of pieces of that package, we will need to be able
to store and retrieve these piece hashes using the P2P protocol. In
addition to storing the file download location in the DHT (which
would still be used for small files), a peer will store a
\emph{torrent string} containing the peer's hashes of the pieces of
the larger files. These piece hashes are compared ahead of time
to determine which peers have the same piece hashes (they all
should), and then used during the download to verify the pieces of
the downloaded package.

@@

We have created a sample implementation that functions as described
in section \ref{opportunity}, and is freely available for other
distributors to download and modify \cite{apt-p2p}. This software,
called \texttt{apt-p2p}, interacts with the \texttt{apt} tool which
is found in most Debian-based Linux distributions. \texttt{apt} uses
SHA1 hashes to verify most downloaded files, including the large
index files that contain the hashes of the individual packages. We
chose this distribution system as it is familiar to us, it allows
software contributions, and there are interesting statistics
available for analyzing the popularity of the software packages
\cite{popcon}.

Since all requests from \texttt{apt} are in the form of HTTP downloads from a
server, the implementation takes the form of a caching HTTP proxy.
Making a standard \texttt{apt} implementation use the proxy is then
as simple as prepending the proxy location and port to the front of
the mirror name in \texttt{apt}'s configuration file (i.e.
``http://localhost:9977/mirrorname.debian.org/\ldots'').

We created a customized DHT based on Khashmir \cite{khashmir}, which
is an implementation of Kademlia \cite{kademlia} using methods
familiar to BitTorrent developers. Khashmir is also the same DHT
implementation used by most of the existing BitTorrent clients to
implement trackerless operation. The communication is all handled by
UDP messages, and RPC (remote procedure call) requests and responses
are all \emph{bencoded} in the same way as BitTorrent's
\texttt{.torrent} files. Khashmir uses the high-level Twisted
event-driven networking engine \cite{twisted}, so we also use
Twisted in our sample implementation for all other networking needs.
More details of this customized DHT can be found below in
Section~\ref{custom_dht}.

Downloading is accomplished by sending simple HTTP requests to the
peers identified by lookups in the DHT to have the desired file.
Requests for a package are made using the package's hash, properly
encoded, as the URL to request from the peer. The HTTP server used
for the proxy also doubles as the server listening for requests for
downloads from other peers. All peers support HTTP/1.1, both in the
server and the client, which allows for pipelining of multiple
requests to a peer, and the requesting of smaller pieces of a large
file using the HTTP Range request header.

@@

\subsection{DHT Background}
\label{dht}

@@

\subsection{Response Time}
\label{response_time}

@@

\subsection{Multiple Values}
\label{multiple_values}

@@

\subsection{\textbf{OPTIONAL}: BitTorrent's Improvements}
\label{bittorrent_dht}

In the several years that some BitTorrent clients have been using a
Kademlia-based DHT for tracker-less operation, the developers have
made many enhancements which we can take advantage of. One of the
most important is a security feature added to stop malicious nodes
from subscribing other nodes as downloaders. When a node issues a
request to another node to store its download info for that key, it
must include a \emph{token} returned by the storing node in a recent
\texttt{find\_node} request. The token is created by hashing the IP
address of the requesting peer with a temporary secret that expires
after several minutes. This prevents the requesting peer from faking
its IP address in the store request, since it must first receive a
response from a \texttt{find\_node} on that IP.

We also made some BitTorrent-inspired changes to the parameters of
the DHT originally specified by the authors of Kademlia. Since, as
we will show later, peers stay online in our system for much longer
periods of time, we reduced Kademlia's \emph{k} value from 20 to 8.
The value is supposed to be large enough such that any given
\emph{k} nodes are unlikely to fail within an hour of each other,
which is very unlikely in our system given the long uptimes of
nodes (see Figure~\ref{duration_online_1} in Section~\ref{analysis}.
We also increased the number of concurrent outstanding
requests allowed from 3 to 6 to speed up the recursive key finding
processes.

\subsection{\textbf{OPTIONAL}: Other Changes}
\label{other_changes}

We added one other new RPC request that nodes can make:
\texttt{join}. This request is only sent on first loading the DHT,
and is usually only sent to the bootstrap nodes that are listed for
the DHT. These bootstrap nodes will respond to the request with the
requesting peer's IP and port, so that the peer can determine what
its oustide IP address is and whether port translation is being
used. In the future, we hope to add functionality similar to STUN
\cite{STUN}, so that nodes can detect whether they are NATted and
take appropriate steps to circumvent it.

In addition, we have allowed peers to store values in the DHT, even
if the hash they are using is not the correct length. Most of the
keys used in the DHT are based on the SHA1 hash, and so they are 20
bytes in length. However, some files in the targetted Debian package
system are only hashed using the MD5 algorithm, and so the hashes
retrieved from the server are 16 bytes in length. The DHT will still
store values using these keys by padding the end of them with zeroes
to the proper length, as the loss of uniqueness from the 4 less
bytes is not an issue. Also, the DHT will happily store hashes that
are longer than 20 bytes, such as those from the 32 byte SHA256
algorithm, by truncating them to the correct length. Since the
requesting peer will have the full length of the hash, this will not
affect its attempt to verify the downloaded file.

@@

We also determined which users are
behind a firewall or NAT, which is one of the main problems of
implementing a peer-to-peer network. These peers will be
unresponsive to DHT requests from peers they have not contacted
recently, which will cause the peer to wait for a timeout to occur
(currently 9 seconds) before moving on. They will also be
unable to contribute any upload bandwidth to other peers, as all
requests for packages from them will also timeout. From
Figure~\ref{walker_peers}, we see that approximately half of all
peers suffered from this restriction.

@@

Figure~\ref{duration_peers} shows the cumulative distribution of how
long a connection from a peer can be expected to last. Due to our
software being installed as a daemon that is started by default
every time their computer boots up, peers are expected to stay for a
long period in the system.
Indeed, we find that 50\% of connections last longer than 5
hours, and 20\% last longer than 10 hours. These connections are
much longer than those reported by Saroiu et. al. \cite{saroiu2001}
for other P2P systems, which had 50\% of Napster and Gnutella
sessions lasting only 1 hour.

@@

\textbf{OPTIONAL}: Since our peers are much longer-lived than other P2P systems, we
also looked at the fraction of peers that stay online for another 6
hours. Figure~\ref{duration_online_6} shows that over 60\% of peers
that are online for 10 hours will stay online for another 6.
However, we see an interesting decrease in this fraction between 8
and 12 hours, which can also be seen in
Figure~\ref{duration_online_1}. We believe this to be due to desktop
users, who regularly turn off their computers at night.

@@

\textbf{OPTIONAL}: Others have also used DHTs to support this type of functionality.
Kenosis \cite{kenosis} is a P2P Remote Procedure Call
client also based on the Kademlia DHT, but it is meant as a P2P
primitive system on which other tools can be built, and so it has no
file sharing functionality at all. Many have used a DHT as a drop-in
replacement for the tracker in a BitTorrent system
\cite{bittorrent-dht, azureus-dht}, but such systems only use the
DHT to find peers for the same torrent, so the file sharing uses
traditional BitTorrent and so is not ideal for the reasons listed
in Section~\ref{bittorrent}.

@@

We have designed a generally applicable peer-to-peer content
distribution model to be used by many of the free software
distributors operating today. It makes use of already existing
features of most package management systems to create a
peer-to-peer distribution that should substantially reduce the costs
of hosting the software packages.

We have also implemented our design in freely available software to
be used in conjuction with Debian-based distribution of Linux
software packages. It is currently in use by some users of the
Debian project's distribution, and so serves as an example of the
possibilities that exist.

We feel that our P2P software package distribution model is
well-suited to be used by many free software distributors. We hope
to convince them to adopt such a model in their distribution, or we
may port our existing system to some of the other groups for them to
try.

\textbf{OPTIONAL}: One aspect missing from our model is the removal of old packages
from the cache. Since our implementation is still relatively young,
we have not had to deal with the problems of a growing cache of
obsolete packages consuming all of a user's hard drive. We plan to
implement some form of least recently used (LRU) cache removal
technique, in which packages that are no longer available on the
server, no longer requested by peers, or simply are the oldest in
the cache, will be removed.

The most significant area of improvement still needed in our sample
implementation is to further speed up some of the slower recursive
DHT requests. We hope to accomplish this by further tuning the
parameters of our current system, better exclusion of NATted peers
from the routing tables, and through the use of STUN \cite{STUN} to
circumvent the NATs of the 50\% of the peers that have not
configured port forwarding.

>>>>

\begin{abstract}
The Internet has become an cost-effective
vehicle for software development and release, particular in the free software society.
Given the free nature of these software, there are often a number of users
motivated by altruism to help out with the distribution, so as to promote the healthy development
of this voluntary society. It is thus naturally expected that peer-to-peer distribution can be implemented,
which scale well with large user bases and can easily explore the network resources made available by
the volunteers.

Unfortunately, this application scenario has many unique characteristics, which
make a straightforward adoption of such existing peer-to-peer systems for file sharing as BitTorrent suboptimal. In particular,
a software release often consists of a large number of packages, which are difficult to be distributed individually, but the archive is
too large to be distributed in its entirety. The packages are also being constantly
updated by the loosely-managed developed; yet the interest in a particular version of a package can be very
limited depending on the computer platforms and operating systems used.

In this paper, we propose a novel peer-to-peer assisted distributor design that
addresses the above challenges. It enhances the existing distribution systems by providing compatible and yet more efficient downloading and updating services
for software packages. Our design leads to \texttt{apt-p2p}, a practical implementation that extends the popular \texttt{apt} distributor.  \texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
software packages and will also be available in the next release of Ubuntu. We have addressed the key design issues in \texttt{apt-p2p}, including indexing table customization,
response time reduction, and multi-value extension. They together ensure
that the altruistic users' resources be effectively utilized and thus significantly reduce the currently
sheer bandwidth requirements of hosting the software, as confirmed by our existing real user statistics.
\end{abstract}

@@

With the widespread penetration of broadband accesses, the Internet has become an cost-effective
vehicle for software development and release. This is particularly true for the free software
society whose developers and users are worldwide distributed and work asynchronously. The ever increasing power of
modern programming languages, computer platforms, and operating systems has made these software extremely large and complex,
which are often divided into a huge number of small packages.
Together with their popularity among users,
an efficient and reliable management and distribution of these software over the Internet has become a challenging task.

The existing distributor for free software are mostly based on the client/server model, e.g.,
the Advanced Package Tool (\texttt{apt}) for Linux, which suffers from the well-known bottleneck problem.
Given the free nature of these software, there are often a number of users
motivated by altruism to help out with the distribution, so as to promote the healthy development
of this voluntary society.
We thus naturally expect that peer-to-peer distribution can be implemented in
this context, which scale well with large user bases and can easily explore the network resources made available by
the volunteers.

Unfortunately, this application scenario has many unique characteristics, which
make a straightforward adoption of such existing peer-to-peer systems for file sharing as BitTorrent suboptimal. In particular,
there are too many packages to distribute each individually, but the archive is
too large to distribute in its entirety. The packages are also being constantly
updated by the loosely-managed developed; yet the interest in a particular version of a package can be very
limited. They together make it very difficult to efficiently create and manage torrents/trackers. The random downloading nature of BitTorrent-like systems is also different from the
sequential order used in existing software package distributors. This in turn suppresses interaction with users
given the difficulty in tracking speed and downloading progress.

In this paper, we propose a novel peer-to-peer assisted distributor design that
addresses the above challenges. It enhances the existing distribution systems by providing compatible and yet more efficient downloading and updating services
for software packages. Our design leads to the development of \texttt{apt-p2p}, a practical implementation based on the Debian \footnote{Debian - The Universal Operating System: {\it http://www.debian.org/}} package
distribution system.  We have addressed the key design issues in \texttt{apt-p2p}, including indexing table customization,
response time reduction, and multi-value extension. They together ensure
that the altruistic users' resources be effectively utilized and thus significantly reduce the currently
sheer bandwidth requirements of hosting the software.

\texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
software packages and will also be available in the next release of Ubuntu. We have evaluated our currently deployment  to
determine how effective it is at meeting our goals, and to see what
effect it is having on the Debian package distribution system. In particular, our existing real user statistics
have suggested that it responsively interacts with clients and substantially reduces server cost.

@@

Given the relatively long time for software package downloading, existing package management systems generally display
some kind of indication of speed and completeness for users to
monitor. Since previous client-server downloads occurred in a sequential
fashion, the package management software also measures the speed
based on sequential downloading. To offer comparable user experience, it is naturally to expect that
the new peer-to-peer solution be reasonably responsive at retrieving packages,
preferably enabling sequential downloading order, too.

@@

On the other hand, there are aspects of BitTorrent that are no
longer critical. Specifically, with altruistic peers
and all files being available without uploading, incentives to share
become a less important issue. Also, the availability of seeds are
not critical either, as the mirror sites serve in that capacity
already.

@@

We now present the design of our peer-to-peer assisted distributor for free software package releases and
updates. A key principle in our design is that the new functionalities implemented in our distributor should be transparent to users,
thus offering the same experience as using conventional software management systems, despite enhanced efficiency.
That said, we assume that the user is still attempting to download packages from a
server, but the requests will be proxied by our peer-to-peer program. The server is always available and has all of the package files.
In addition, the cryptographic hash of the packages will be available
separately from the package itself, and is usually contained in an
\emph{index} file which also contains all the packages' names,
locations and sizes.

@@

Our model for using peer-to-peer to enhance package distribution is shown in
Figure~\ref{model}. As shown in Phase~1, our program will act as a
proxy (1,2), downloading (3) and caching all files communicated
between the user and the server (4). It will therefore also have
available the index files containing the cryptographic hashes all
packages. Later, in Phase~2, upon receiving a request from the user
to download a package (5), our program will search the index files
for the package being requested and find its hash (6). This hash can
then be looked up recursively in an indexing structure (a Distributed Hash Table [] in our implementation) (7), which will return a
list of peers that have the package already (8). The package can
then be downloaded from the peers (11,12), it can be verified using
the hash (13), and if valid can be returned to the user (14). The
current nodes location is also added to the DHT for that hash (15),
as it is now a source for others to download from.

@@

There are several ways to implement the desired peer-to-peer functionality
into the existing package management software. The functionality can
be directly integrated through modifications to the software, though
this could be difficult as the peer-to-peer functionality should be running
at all times. This is needed both for efficient lookups and to
support uploading of already downloaded packages. Unfortunately, the
package management tools typically only run until the download and
install request is complete.

Many of the package management software implementations use HTTP
requests from web servers to download the packages, which makes it
possible to implement the peer-to-peer aspect as an almost standard HTTP
caching proxy. This proxy will run as a daemon in the background,
listening for requests from the package management tool for package
files, as well as serving (uploading) cached package to other peers.
It will get uncached requests first from the peer-to-peer system, or
fall back to the normal HTTP request from a server should it not
be found. For methods that don't use HTTP requests, other types of
proxies may also be possible.

@@

Although not necessary, we recommend implementing a download
protocol that is similar to the protocol used to fetch packages from
the distributor's servers. This simplifies the peer-to-peer program, as it
can then treat peers and mirrors almost identically when requesting
packages. In fact, the mirrors can be used when there are only a few
slow peers available for a file to help speed up the download
process.

@@

Since the package management system only stores a hash of the entire
package, and not of pieces of that package, we will need to be able
to store and retrieve these piece hashes using the peer-to-peer protocol. In
addition to storing the file download location in the DHT (which
would still be used for small files), a peer will store a
\emph{torrent string} containing the peer's hashes of the pieces of
the larger files. These piece hashes are compared ahead of time
to determine which peers have the same piece hashes (they all
should), and then used during the download to verify the pieces of
the downloaded package.

@@

We have created a sample implementation that functions as described
in section \ref{opportunity}, and is freely available for other
distributors to download and modify \cite{apt-p2p}. This software,
called \texttt{apt-p2p}, interacts with the popular \texttt{apt} tool. This tool 
is found in most Debian-based Linux distributions, with related statistics
available for analyzing the popularity of the software packages
\cite{popcon}.

Since all requests from \texttt{apt} are in the form of HTTP downloads from a
server, our implementation takes the form of a caching HTTP proxy.
Making a standard \texttt{apt} implementation use the proxy is then
as simple as prepending the proxy location and port to the front of
the mirror name in \texttt{apt}'s configuration file (i.e.
``http://localhost:9977/mirrorname.debian.org/\ldots'').

We created a customized DHT based on Khashmir \cite{khashmir}, which
is an implementation of Kademlia \cite{kademlia}. Khashmir is also the same DHT
implementation used by most of the existing BitTorrent clients to
implement trackerless operation. The communication is all handled by
UDP messages, and RPC (remote procedure call) requests and responses
are all \emph{bencoded} in the same way as BitTorrent's
\texttt{.torrent} files. Khashmir uses the high-level Twisted
event-driven networking engine \cite{twisted}, so we also use
Twisted in our sample implementation for all other networking needs.
More details of this customized DHT can be found below in
Section~\ref{custom_dht}.

Downloading is accomplished by sending simple HTTP requests to the
peers identified by lookups in the DHT to have the desired file.
Requests for a package are made using the package's hash, properly
encoded, as the URL to request from the peer. The HTTP server used
for the proxy also doubles as the server listening for requests for
downloads from other peers. All peers support HTTP/1.1, both in the
server and the client, which allows for pipelining of multiple
requests to a peer, and the requesting of smaller pieces of a large
file using the HTTP Range request header. Like in \texttt{apt}, 
SHA1 hashes are then used to verify downloaded files, including the large
index files that contain the hashes of the individual packages.

@@

\subsection{Kademlia DHT Background}
\label{dht}

@@

\subsection{Response Time Optimization}
\label{response_time}

@@

\subsection{Extension with Multiple Values}
\label{multiple_values}

@@

@@

We also determined which users are
behind a firewall or NAT, which is one of the main problems of
implementing a peer-to-peer network. These peers will be
unresponsive to DHT requests from peers they have not contacted
recently, which will cause the peer to wait for a timeout to occur
(currently 9 seconds) before moving on. They will also be
unable to contribute any upload bandwidth to other peers, as all
requests for packages from them will also timeout. From
Figure~\ref{walker_peers}, we see that approximately half of all
peers suffered from this restriction. To address this problem, we added one other new RPC request that nodes can make:
\texttt{join}. This request is only sent on first loading the DHT,
and is usually only sent to the bootstrap nodes that are listed for
the DHT. These bootstrap nodes will respond to the request with the
requesting peer's IP and port, so that the peer can determine what
its oustide IP address is and whether port translation is being
used. In the future, we hope to add functionality similar to STUN
\cite{STUN}, so that nodes can detect whether they are NATted and
take appropriate steps to circumvent it.

@@

Figure~\ref{duration_peers} shows the cumulative distribution of how
long a connection from a peer can be expected to last. Due to our
software being installed as a daemon that is started by default
every time their computer boots up, peers are expected to stay for a
long period in the system.
Indeed, we find that 50\% of connections last longer than 5
hours, and 20\% last longer than 10 hours. These connections are
much longer than those reported by Saroiu et. al. \cite{saroiu2001}
for other peer-to-peer systems, which had 50\% of Napster and Gnutella
sessions lasting only 1 hour.

@@

\textbf{OPTIONAL}: Since our peers are much longer-lived than other peer-to-peer systems, we
also looked at the fraction of peers that stay online for another 6
hours. Figure~\ref{duration_online_6} shows that over 60\% of peers
that are online for 10 hours will stay online for another 6.
However, we see an interesting decrease in this fraction between 8
and 12 hours, which can also be seen in
Figure~\ref{duration_online_1}. We believe this to be due to desktop
users, who regularly turn off their computers at night.

@@

\textbf{OPTIONAL}: Others have also used DHTs to support this type of functionality.
Kenosis \cite{kenosis} is a peer-to-peer Remote Procedure Call
client also based on the Kademlia DHT, but it is meant as a peer-to-peer
primitive system on which other tools can be built, and so it has no
file sharing functionality at all. Many have used a DHT as a drop-in
replacement for the tracker in a BitTorrent system
\cite{bittorrent-dht, azureus-dht}, but such systems only use the
DHT to find peers for the same torrent, so the file sharing uses
traditional BitTorrent and so is not ideal for the reasons listed
in Section~\ref{bittorrent}.

@@

In this paper, we have provided strong evidence that free software package distribution 
and update exhibit many distinct characteristics, which call for new designs other
than the existing peer-to-peer systems for file sharing. To this end, we have 
present \texttt{apt-p2p}, a novel peer-to-peer distributor that sits between
client and server, providing efficient and transparent downloading and updating services
for software packages. We have addressed the key design issues in \texttt{apt-p2p}, including DHT customization,
response time reduction, and multi-value extension. \texttt{apt-p2p}  has been used in conjuction with Debian-based distribution of Linux
software packages and will also be available in the next release of Ubuntu. Existing real user statistics
have suggested that it interacts well with clients and substantially reduces server cost.

There are many future avenues toward improving our implementation. Besides 
evaluating its performance in larger scales, we are particularly interest in further speeding up some of the slower recursive
DHT requests. We expect to accomplish this by fine tuning the
parameters of our current system, better exclusion of NATted peers
from the routing tables, and through the use of STUN \cite{STUN} to
circumvent the NATs of the 50\% of the peers that have not
configured port forwarding.

One aspect missing from our model is the removal of old packages
from the cache. Since our implementation is still relatively young,
we have not had to deal with the problems of a growing cache of
obsolete packages consuming all of a user's hard drive. We plan to
implement some form of least recently used (LRU) cache removal
technique, in which packages that are no longer available on the
server, no longer requested by peers, or simply are the oldest in
the cache, will be removed.
